---
title: "appendix_b"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{appendix_b}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Storage and formats of model input and output

Depending on the resolution of the model and the size of the country, the size of the input (raw and processed) and output files can become quite large. Hence it is important to ensure sufficient disk space to store the files. To give you an idea, all files related to the China example have a total size of [@ADD], while the total size for the Malawi example is [@ADD]. We also encountered some problems when performing some of the GDAL operations using DropBox, probably related to the  synchronization of files. If you encounter problems when saving to Dropbox, try saving to a folder that is not being  synchronizated automatically.  

In most cases we use the following file formats to store data:
- Raster maps are stored using Geotiff (*.tif).
- Shape files (i.e. vector/polygon data) are stored in the R .rds data format, which ensures the data structure is maintained and data can be quickly loaded into R. The shapefile with ADM information is also saved in .shp format as this is needed for GDAL.
- Data files are stored in the R .rds data format. For files with lists of information (e.g. list of crops and ADMs in the country) we use .csv files for quick referencing (e.g. using Excel).
- GAMS code is saved in .gms [@UPDATE]
- GAMS data files are stored in .gdx [@UPDATE]


## Predefined dimensions

### Number of crops

SPAMc models 40 different crop (and crop groups) that together cover the full agricultural sector and are each identified by a four letter code (Table 1).^[Global SPAM uses 42 different crops because it includes two types of millet and two types of coffee [@Wood-Sichra2016]. As this level of detail is not supported by most national statistics and FAOSTAT, the have been merged into millet and coffee, respectively.] The main reason for this classification is the limited availability of crop-specific biophysical suitability maps, which form a key element in the crop allocation process (see [spatial data](process_spatial_data.html) for more information). It would be relatively easy to add new crops by splitting them off from broader crop groups  (e.g. tropical and temperate fruits, and vegetables) if appropriate agricultural statistics and suitability maps would be available. We plan to add an example on how to do this in future updates. The actual number of crops in the model is determined by the number of crops that are actually grown in the target country and identified by the availability of national statistics.

```{r}
# Add 40 crop table
```


### Number of farming systems

Also similar to global SPAM, SPAMc distinguises between four different farming systems:

- Rainfed subsistence
- Rainfed low-input
- Rainfed high-input
- Irrigated


