---
title: "Model setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
  bib/mapspam.bib
bibliography: bib/mapspam.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
  
# Presetup as it cannot be done in relevant chunk because that one is not evaluated.
library("mapspam2globiom")
spamc_path <- "C:/Users/dijk158/Dropbox/mapspam2globiom_mwi"
```

To setup SPAMc two steps need to be taken (the order is important!). It is assumed that all required software is already installed and working (see [software and hardware](software.html)). As [mentioned](template.html), it is by far the easiest to install the template in an RStudio project first and then start to work on your own country SPAMc version as all the steps discussed from now on are already pre-coded in the template.

## Create SPAMc folder structure

`mapspam2globiom` includes a function to create all necessary folders `create_spam_folders()`. The only input it requires is a path on your harddisk. We do not recommend using a server to store the model as the potential increase in file access time might seriously affect the time it takes to process some of the spatial datasets which can be very large. Note that it is perfectly fine to store large *raw data* files on a server (see below). 

```{r eval=FALSE}
# Load mapspam2globiom
library("mapspam2globiom")

# Set the folder where the model will be stored
# Note that R uses forward slashes even in Windows!!
spamc_path <- "C:/Users/dijk158/Dropbox/mapspam2globiom_mwi"

# Create SPAMc folder structure in the spamc_path
create_spam_folders(spamc_path)
```


## Set model parameters and location of files

Nearly all functions in `mapspam2globiom` need input on (1) key parameters that determine the design of the model and how will be solved and (2) the location of raw and intermediate data, which will be processed (see [model structure](model_structure.html). Both pieces of information are bundled and stored by the function `set_spam_par()` in a `spam_par` object. 

```{r}
# Set SPAMc parameters
par <- spam_par(spam_path = spamc_path,
 iso3c = "MWI",
 year = 2010,
 res = "5min",
 adm_level = 1,
 solve_level = 0,
 model = "max_score")

# check class
class(par)

# Show parameters
print(par)
```

This object need to be saved in the `parameters` folder that was created together with all the other folders when running `create_spam_folders`. Elements in a `spam_par` object class can be extracted by using the dollar sign ($). This feature is used to save `par` the temporary object of class `spam_par` into the relevant folder.

```{r eval=FALSE}
# Save 
saveRDS(par, file.path(par$spam_path, "parameters/spam_par.rds"))
```

That is all it takes to set up the SPAMc model! The next step is processing the [raw subnational statistics](process_subnat_stat.html) and [spatial data](process_spatial_data.html).
