---
title: "Model setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
  bib/mapspam.bib
bibliography: bib/mapspam.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
```

To setup SPAMc several steps need to be taken. By now, we assume that all required software is [installed](installation.html) and working. It is by far the easiest to create a new RStudio project and use the mapspam2globiom_mwi for Malawi [template](template.rmd) as a basis of your country SPAMc version. All the steps described in the Run SPAMc section use the code in mapspam2globiom_mwi repository as an example code so there is no need to start from scratch. All scripts related to model setup can be found in the `01_model_setup` folder of the template. The most important is that you change the location on your hard disk where the model will be stored and set the model parameters (see below), which is done in the script `01_model_setup.r` in the template. As the model parameters are needed as input in nearly all functions, this particular script is always loaded at the beginning of all the other scripts using the the `source()` command. In this way, you only have to setup up the model once. It also ensures all the necessary packages are loaded into R. 


## Set model parameters and location of files

Nearly all functions in `mapspam2globiom` need input on (1) key parameters that determine the design of the model and how will be solved and (2) the location of raw and intermediate data, which will be further processed. Both pieces of information are bundled and stored by the function `set_spam_par()` in a `spam_par` object. The example below does this for Malawi and stores the parameters in `param`. Param is an arbitrary name but we recommend not to change it as we use `param` as input in nearly all functions that are described next. One piece of information that is also stored in the `spam_par` object is the coordinate reference system of the maps. This is automatically set to WSG84 (epsg:4326).

```{r}
# Load mapspam2globiom
library("mapspam2globiom")

# Set the folder where the model will be stored
# Note that R uses forward slashes even in Windows!!
spamc_path <- "C:/Users/dijk158/Dropbox/mapspam2globiom_mwi"

# Set SPAMc parameters
param <- spam_par(spam_path = spamc_path,
 iso3c = "MWI",
 year = 2010,
 res = "5min",
 adm_level = 1,
 solve_level = 0,
 model = "max_score")

# Show parameters
print(param)
```


## Create SPAMc folder structure

`mapspam2globiom` includes a function to create all necessary folders `create_spam_folders()`. The only input it requires is the `param` which contains all the model parameters. The function will create three folders (1) `raw_data` is created in the main `spam_path` unless indicated otherwise by the user and serves to store all the raw input data, (2) `processed_data` will be created in the main `spam_path` and is used to save all the intermediary data products as well as the final results and (3) `mappings` will be created in the main `spam_path` and contains all the lists and mappings that are needed to define the model (e.g. number crops) and harmonize various data sources (concordance tables). All tables are in csv format and will be copied automatically into the `mappings` folder. Note that you can adjust as some of these tables if needed (e.g. adding a new crop), they are only copied if the files do not exist. If for some reason you want to replace or regenerate them, simply remove the folder and it will be created again when `create_spam_folders` is run. 

```{r eval=FALSE}
# Create SPAMc folder structure in the spamc_path
create_spam_folders(param)
```


## Prepare subnational administrative unit map

To allocate the subnational crop area information, a complementary map (most likely in the form of a shapefile) with the location of the subnational administrative units is needed. The processing of this map cannot be captured by a single function as it involves several decisions that need to be made by the user. These are explained next. 

Before the shapefile can be processed you need to load it into R by setting the name (in this case `adm_2010_MWI.shp`), which is stored in the `raw_data/adm/` folder. The `st_transform` command projects the map to WSG84 so that it is consistent with all the other spatial data. 

```{r eval = FALSE}
# replace the name of the shapefile with that of your own country.
iso3c_shp <- "adm_2010_MWI.shp"

# load shapefile
adm_map_raw <- read_sf(file.path(param$spam_path, glue("raw_data/adm/{iso3c_shp}")))

# plot
plot(adm_map_raw$geometry)

# Project to standard global projection
adm_map <- adm_map_raw %>%
  st_transform(param$crs)
```

For the model to work, it is essential that the attribute table with the names and codes of the administrative units has the correct column names. To make this easier, you only have to add the current header of the names and the codem, which will be stored in `admX_name_orig` and `admX_code_orig`, respectively, where X is the administrative unit level. If you only have data at level 0 and 1, simply delete the lines of script that refer to administrative unit level 2, etc. In case of Malawi, we have data at level 0-1 so all lines of script are relevant. Columns in the attribute table that are not relevent will be automatically deleted. We also union all polygons with the same name and code. Hence, if you only have statistics for the combination of two or more subnational regions, you can simply give them the same name and code and they will automatically be merged.

```{r eval = FALSE}
# Check names
head(adm_map)
names(adm_map)

# Set the original names, i.e. the ones that will be replaced. Remove adm1
# and/or adm2 entries if such data is not available.
adm0_name_orig <- "ADM0_NAME"
adm0_code_orig <- "FIPS0"
adm1_name_orig <- "ADM1_NAME"
adm1_code_orig <- "FIPS1"
adm2_name_orig <- "ADM2_NAME"
adm2_code_orig <- "FIPS2"

# Replace the names
names(adm_map)[names(adm_map) == adm0_name_orig] <- "adm0_name"
names(adm_map)[names(adm_map) == adm0_code_orig] <- "adm0_code"
names(adm_map)[names(adm_map) == adm1_name_orig] <- "adm1_name"
names(adm_map)[names(adm_map) == adm1_code_orig] <- "adm1_code"
names(adm_map)[names(adm_map) == adm2_name_orig] <- "adm2_name"
names(adm_map)[names(adm_map) == adm2_code_orig] <- "adm2_code"

# Only select relevant columns
adm_map <- adm_map %>%
  dplyr::select(adm0_name, adm0_code, adm1_name, adm1_code, adm2_name, adm2_code)

# Union separate polygons that belong to the same adm    
adm_map <- adm_map %>%
  group_by(adm0_name, adm0_code, adm1_name, adm1_code, adm2_name, adm2_code) %>%
  summarize() %>%
  ungroup() %>%
  mutate(adm0_name = param$country,
         adm0_code = param$iso3c)

# Check names
head(adm_map)
names(adm_map)
```

Polygons of administrative units where no crops should be allocated must be removed. If this is not done, in rare cases when the area statistics are larger than the cropland exent, the model might still allocate crops in these regions.^[@ADD].The Malawi polygon includes two of such regions. One is the Area under National Administration, which is the part of Lake Malawi that belongs to Malawi and Likoma, several small islands in the lake that are not covered by the statistics. The script below shows how to remove them. It there is no reason to remove polygons, delete the lines of script. By running `create_adm_map_pdf()` a pdf file with maps of the subnational regions is created in the `processed_data/maps/adm`  folder. Finally, you need to save the maps in shapefile and rds format in the relevant folder.

```{r eval = FALSE}
par(mfrow=c(1,2))
plot(adm_map$geometry, main = "ADM all polygons")

# Set names of ADMs that need to be removed from the polygon. 
adm1_to_remove <- c("Area under National Administration")
adm2_to_remove <- c("Likoma")

# Remove ADMs
adm_map <- adm_map %>%
  filter(adm1_name != adm1_to_remove) %>%
  filter(adm2_name != adm2_to_remove)

plot(adm_map$geometry, main = "ADM polygons removed")
par(mfrow=c(1,1))

# Create pdf file with the location of administrative units
create_adm_map_pdf(param)

# Save adm maps
temp_path <- file.path(param$spam_path, glue("processed_data/maps/adm/{param$res}"))
dir.create(temp_path, showWarnings = FALSE, recursive = TRUE)

saveRDS(adm_map, file.path(temp_path, glue("adm_map_{param$year}_{param$iso3c}.rds")))
write_sf(adm_map, file.path(temp_path, glue("adm_map_{param$year}_{param$iso3c}.shp")))
```

To link the subnational statistics to other spatial data a table is needed (`adm_list`) that contains information on how the administrative units at various levels are nested, which is exactly the information that is stored in the attribute table of the administrative unit map. The following code, extracts the list from the shapefile and saves it as csv file in the `processed_data/lists` folder. 

```{r eval = FALSE}
# Create adm_list
create_adm_list(adm_map, param)
```


## Create grid and rasterize subnational administrative unit map

The package comes with `create_grid()` a function that creates a spatial grid at selected spatial resolution (i.e. 30 arcsec or 5 arcmin). The grid is subsequently used by `rasterize_adm_map()`, which rasterizes the administrative unit map at the selected resolution. Both the grid and the rasterized map are required by the model as input.

```{r eval = FALSE}
# Create grid
create_grid(param)

# Rasterize administrative unit map
rasterize_adm_map(param)
```

This is all it takes to set up the SPAMc model! The next step is processing the [raw subnational statistics](process_subnational_statistics.html) and [spatial data](process_spatial_data.html).
