---
title: "Post processing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Post processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
resource_files:
  bib/mapspam.bib
bibliography: bib/mapspam.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Inspect results and Create maps 

View results

```{r eval = FALSE}
view_4panel
```



```{r eval = FALSE}
create_all_tif(param)
```



## Replacing land cover and land use information in GLOBIOM

TODO, add coffee as additional crop!!!!


With respect to the distribution of crops, GLOBIOM uses two related data layers. A table with area information on the land cover per simu, distinguishing six different land use classes:

- Total agricultural land (CrpLnd, OthAgri)
- Grassland (Grass)
- Forest (Forest)
- Wetlands (WetLnd)
- Other natural vegetation (OthNatLnd)
- Not relevant land cover (NotRel)

And a corresponding table with more detailed information on the crop area per simu. The land cover class Total agricultural area is divided into two parts: (1) Cropland (CrpLnd), which presents the location of the 20 crops that are modelled by GLOBIOM and (2) Other agricultural land, which presents the location of all the other crops.

More information on the construction of the land cover and land use layers can be found in @Skalsky

Two files are needed to update the land use and land cover maps in GLOBIOM:

* A new land cover map that shows the location of 
*

`create_globiom_input(param)` 

```{r eval = FALSE}
create_globiom_input(param)
```




## Change GLOBIOM
